systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2006-02-17 paper/trunk/nlsystemfit.tex
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31

where, $r$ is a column vector for the residuals for each equation,
%% what is the difference to \epsilon?
$X$ is matrix of the partial derivatives of the dependent variable 
with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
%% is this correct?
%% in the linear section X is used for the regressors, 
%% thus we should use a different variable name here.
$W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
matrix of the instrument variables, and $I$ is an $n \times n $
identity matrix and $S$ is the estimated variance-covariance matrix between the
equations

@@

\textbf{The residual variance-covariance matrix can be calculated in 
different ways (section~\ref{sec:residcov}). 
It should be relatively easy to implement this also in nlsystemfit(), 
e.g. the function 'calcRCov' in systemfit() could be moved outside
systemfit() that it can be used also by nlsystemfit().}

>>>>

where, $r$ is a column vector for the residuals for each equation,
%% what is the difference to \epsilon?
$X$ is matrix of the partial derivatives of the dependent variable 
with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
%% is this correct?
%% in the linear section X is used for the regressors, 
%% thus we should use a different variable name here.
$W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
matrix of the instrument variables, and $I$ is an $n \times n $
identity matrix and $S$ is the estimated covariance matrix between the
equations

@@

\textbf{The residual covariance matrix can be calculated in
different ways (section~\ref{sec:residcov}). 
It should be relatively easy to implement this also in nlsystemfit(), 
e.g. the function 'calcRCov' in systemfit() could be moved outside
systemfit() that it can be used also by nlsystemfit().}

### 2006-02-17 paper/trunk/systemfit_estimation.tex
added description of methods "WSUR" and "W3SLS"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@599 591db2bd-e0a0-4d22-8c19-2192dbb7da31
substituted "covariance matrix" for "variance-covariance matrix" and "\Cov [ . ]" for "Var [ . ]", where "\Cov" is a command for "\text{Cov}"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@605 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary least squares (OLS)}

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X'X \right)^{-1} X'y
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
   \label{eq:olsVarSameSigma}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations
are allowed to have different variances,
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
   \label{eq:olsVarSingleSigma}
\end{equation}
with $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

If no cross-equation parameter restrictions are imposed, the simultaneous 
OLS estimation of the system leads to the same parameter estimates 
as an equation-wise OLS estimation.
The variance-covariance matrix of the parameters from an equation-wise
OLS estimation is equal to the variance-covariance matrix obtained
by equation (\ref{eq:olsVarSingleSigma}).

@@

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.
Like the OLS estimates these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

The SUR estimator can be obtained by:%
\footnote{
To calculate $\Omega^{-1}$ it is not necessary to invert 
the (huge) $\Omega$ matrix, 
because it is sufficient to invert the (small) $\Sigma$ matrix and
calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$
\citep[p.\ 342]{greene03}.
}
\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.
And the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}

@@

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
\end{equation}
At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y 
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1}
   \label{eq:2slsVarSameSigma}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
If the disturbance terms of the individual equations
are allowed to have different variances, 
the variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
   \label{eq:2slsVarSingleSigma}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.
The variance-covariance matrix of the estimated parameters is
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
\end{equation}

@@

The standard 3SLS estimator can be obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
   \label{eq:3slsGls}
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.
Its variance-covariance matrix is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   \widehat{X} \right)^{-1}
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( E \left[ u_i | H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( E \left[ u | H \right]) = 0 )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right]
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:

One of these approaches is based on instrumental variable estimation
(IV-3SLS):
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
   \label{eq:3slsIv}
\end{equation}
The variance-covariance matrix of this IV-3SLS estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1} 
   X \right)^{-1}
\end{equation}
Another approach is based on the Generalized Method of Moments (GMM)
estimator (GMM-3SLS):
\begin{equation}
   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
   \label{eq:3slsGmm}
\end{equation}
The variance-covariance matrix of the GMM-3SLS estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = 
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y 
   \label{eq:3slsSchmidt}
\end{equation}
The variance-covariance matrix of this estimator is:
\begin{equation}
   Var \left[ \widehat{\beta} \right] = 
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
\end{equation}
The econometrics software EViews uses following approach:
\begin{equation}
   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
   \label{eq:3slsEViews}
\end{equation}
with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
(equation \ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the variance-covariance matrix of the 3SLS estimator.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary least squares (OLS)}
\label{sec:ols}

The Ordinary Least Squares (OLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X'X \right)^{-1} X'y
\end{equation}
These estimates are efficient only if the disturbance terms are not
contemporaneously correlated, which means
$\sigma_{ij} = 0 \; \forall \; i \neq j$.
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \sigma^2 \left( X'X \right)^{-1}
   \label{eq:olsCovSameSigma}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
This assumes that the disturbances of all equations have the
same variance.

If the disturbance terms of the individual equations
are allowed to have different variances,
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
   \label{eq:olsCovSingleSigma}
\end{equation}
with $\Omega = \Sigma \otimes I$,
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

If no cross-equation parameter restrictions are imposed, the simultaneous 
OLS estimation of the system leads to the same parameter estimates 
as an equation-wise OLS estimation.
The covariance matrix of the parameters from an equation-wise
OLS estimation is equal to the covariance matrix obtained
by equation (\ref{eq:olsCovSingleSigma}).

@@

The Weighted Least Squares (WLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.
Like the OLS estimates these estimates are only efficient
if the disturbance terms are not contemporaneously correlated.
The covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}
If no cross-equation parameter restrictions are imposed,
the parameter estimates are equal to the OLS estimates.

@@

The SUR estimator can be obtained by:%
\footnote{
To calculate $\Omega^{-1}$ it is not necessary to invert 
the (huge) $\Omega$ matrix, 
because it is sufficient to invert the (small) $\Sigma$ matrix and
calculate it by $\Omega^{-1} = \Sigma^{-1} \otimes I$
\citep[p.\ 342]{greene03}.
}
\begin{equation}
   \widehat{\beta} = \left( X' \Omega^{-1} X \right)^{-1} X' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.
And the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( X' \Omega^{-1} X \right)^{-1}
\end{equation}

@@

At the first stage new ('fitted') regressors are obtained by
\begin{equation}
   \widehat{X_i} = H_i \left( H_i' H_i \right)^{-1} H_i' X
\end{equation}
At the second stage the unbiased two-stage least squares estimates
of $\beta$ are obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \widehat{X} \right)^{-1} 
   \widehat{X}' y 
   \label{eq:beta2sls}
\end{equation}
If the whole system is treated as one single equation,
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \sigma^2 \left( \widehat{X}'
   \widehat{X} \right)^{-1}
   \label{eq:2slsCovSameSigma}
\end{equation}
with $\sigma^2 = E \left( u' u \right)$.
If the disturbance terms of the individual equations
are allowed to have different variances, 
the covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1}
   \label{eq:2slsCovSingleSigma}
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.

@@

The Weighted Two-Stage Least Squares (W2SLS) estimator of the system 
is obtained by
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
\end{equation}
with $\Omega = \Sigma \otimes I$, 
$\sigma_{ij} = 0 \; \forall \; i \neq j$ and
$\sigma_{ii} = E \left( u_i' u_i \right)$.
The covariance matrix of the estimated parameters is
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1}
\end{equation}

@@

The standard 3SLS estimator can be obtained by:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
   \label{eq:3slsGls}
\end{equation}
with $\Omega = \Sigma \otimes I$ and
$\sigma_{ij} = E \left( u_i' u_j \right)$.
Its covariance matrix is:
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
   \widehat{X} \right)^{-1}
   \label{eq:cov3sls}
\end{equation}
While an unbiased 2SLS or W2SLS estimation requires only that
the instrumental variables and the disturbance terms of each single 
equation are uncorrelated
$( E \left[ u_i | H_i \right]) = 0 \; \forall \; i )$,
\cite{schmidt90} points out that this estimator is only consistent 
if all disturbance terms and all instrumental variables are uncorrelated
$( E \left[ u | H \right]) = 0 )$
with
\begin{equation}
   H =
   \left[ \begin{array}{cccc}
      H_1 & 0 & \cdots & 0\\
      0 & H_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & H_G
   \end{array}\right]
\end{equation}
Since there might be occasions where this cannot be avoided,
\cite{schmidt90} analyses other approaches to obtain 3SLS estimators:

One of these approaches is based on instrumental variable estimation
(IV-3SLS):
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} X 
   \right)^{-1} \widehat{X}' \Omega^{-1} y
   \label{eq:3slsIv}
\end{equation}
The covariance matrix of this IV-3SLS estimator is:
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] = \left( \widehat{X}' \Omega^{-1}
   X \right)^{-1}
\end{equation}
Another approach is based on the Generalized Method of Moments (GMM)
estimator (GMM-3SLS):
\begin{equation}
   \widehat{\beta} = \left( X' H \left( H' \Omega H \right)^{-1}
   H' X \right)^{-1} X' H \left( H' \Omega H \right)^{-1} H' y
   \label{eq:3slsGmm}
\end{equation}
The covariance matrix of the GMM-3SLS estimator is:
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] =
   \left( X' H \left( H' \Omega H \right)^{-1} H' X \right)^{-1}
\end{equation}
A fourth approach developed by \cite{schmidt90} himself is:
\begin{equation}
   \widehat{\beta} = \left( \widehat{X}' \Omega^{-1} \widehat{X} 
   \right)^{-1} \widehat{X}' \Omega^{-1} 
   H \left( H' H \right)^{-1} H' y 
   \label{eq:3slsSchmidt}
\end{equation}
The covariance matrix of this estimator is:
\begin{equation}
   \Cov \left[ \widehat{\beta} \right] =
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} H \left( H' H \right)^{-1} H' \Omega 
   H \left( H' H \right)^{-1} H' \Omega^{-1} \widehat{X}
   \left( \widehat{X}' \Omega^{-1}  \widehat{X} \right)^{-1}
\end{equation}
The econometrics software EViews uses following approach:
\begin{equation}
   \widehat{\beta} = \widehat{\beta}_{2SLS} + 
   \left( \widehat{X}' \Omega^{-1} \widehat{X} \right)^{-1} 
   \widehat{X}' \Omega^{-1} \left( y - X \widehat{\beta}_{2SLS} \right)
   \label{eq:3slsEViews}
\end{equation}
with $\widehat{\beta}_{2SLS}$ is the two-stage least squares estimator
(equation \ref{eq:beta2sls}).
EViews uses the standard 3SLS formula (\ref{eq:cov3sls}) to
calculate the covariance matrix of the 3SLS estimator.
